---
title: "Lab 3: Hypothesis Tests about the Mean."
subtitle: "w203: Statistics for Data Science"
author: "Tako Hisada"
date: "11/08/2017"
output: pdf_document
--- 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

## The Data

The American National Election Studies (ANES) conducts surveys of voters in the United States before and after every presidential election.  You are given a small subset of the 2012 ANES survey, contained in the file ANES\_2012\_sel.csv.

There are a number of special concerns that arise whenever statisticians work with survey data.  In particular, the complete ANES survey data assigns a survey weight to each observation, which corrects for differences in how likely individuals are to be selected, and how likely they are to respond.  For the purposes of this assignment, however, we have removed the survey weights and we ask you to assume that the observations you have are a random sample from the voting population.

For a glimpse into some of the intricacies that go into survey design, take a look at the introduction to the ANES User's Guide and Codebook. 


```{r load}
S = read.csv("ANES_2012_sel.csv")
```

Following is an example of a question asked on the ANES survey:

Where would you place YOURSELF on this scale, or haven't you thought much about this?

Possible answers included:

- 1. Extremely liberal 
- 2. Liberal 
- 3. Slightly liberal 
- 4. Moderate; middle of the road 
- 5. Slightly conservative 
- 6. Conservative 
- 7. Extremely conservative
- -2. Haven't thought much about this 
- -8. Don't know 
- -9. Refused

The variable libcpre\_self records answers before the election, while libcpo\_self records answers after the election.

# Analysis

## 1. Did voters become more liberal or more conservative during the 2012 election?

### Analysis

All survey responses can be forced into numeric values on the scale of 1-7, -2, -8 or -9.

```{r q1_1}
table(S$libcpre_self, as.numeric(S$libcpre_self))
n = length(S$libcpre_self)
```

Our n is `r n`.

In order to measure where people stand on the political spectrum, however, the answers -2, -8 or -9 are not informative. Also because of the presence of these values, the numeric values assigned by R are off by 3 from the numbers present in the survey answer choices (i.e. R assigned 4 to "1. Extremely liberal").

We need to re-assign values to match the survey answer choices and also 0 out the values for the choices -2, -8 and -9.

```{r q1_2}
# Define bins for political standing histograms
bins_standing <- seq(from = 0, to = 7, by = 1)

# Function for generating a vector filled with a specified value 
generateSeq <- function(v, value) {
  seq = c()
  for(i in 1:length(v)) {
    seq[i] = value
  }
  seq
}

# Function for generating political weights nulling invalid answers 
compilePoliticalWeights <- function(v, null_index, skip_zero = FALSE) {
  num_v = as.numeric(v)
  weights = c()
  
  for(i in 1:length(unique(num_v))) {
    if(i <= null_index) {
      if(!skip_zero) {
        weights = c(weights, generateSeq(num_v[num_v == i], 0))        
      }
    } else {
      weights = c(weights, generateSeq(num_v[num_v == i], i - null_index))
    }
  }
  weights
}

libcpre_weights <- compilePoliticalWeights(S$libcpre_self, 3, TRUE)
summary(libcpre_weights)
mu_libcpre = mean(libcpre_weights)
hist(libcpre_weights, main = "Pre-2012 Election Political Standing", 
     xlab = "Political Standing", breaks = bins_standing)
```

Our $\mu$ pre 2012 election is `r mu_libcpre`.

Likewise, we will need to do the same data sanitizing for the post-election data.

```{r q1_3}
table(S$libcpo_self, as.numeric(S$libcpo_self))
```

The data now contains 2 additional invalid values -6 and -7 in addition to -2, -8 and -9 we saw previously.

```{r q1_4}
libcpo_weights <- compilePoliticalWeights(S$libcpo_self, 5, TRUE)
summary(libcpo_weights)
mu_libcpo = mean(libcpo_weights)
hist(libcpo_weights, main = "Post-2012 Election Political Standing", 
     xlab = "Political Standing", breaks = bins_standing)
```

Our $\mu$ post 2012 election is `r mu_libcpo` which is slightly lower than before the election suggesting there were more people idntified with the liberal-side of the political spectrum.

### Hypotheses

From this, we can formulate our hypotheses as below:

$\mu_1$ = Pre-2012 election political standing\newline
$\mu_2$ = Post-2012 election political standing

$$H_0: \mu_1 - \mu_2 = 0$$
$$H_1: \mu_1 - \mu_2 > 0$$

### Statistical Significance

#### Justificaition

Our distribution is fairly normal and our n is sufficiently large. We can conduct a t-test in determining if the post-election mean is in fact lower than the pre-election.

```{r q1_5}
t.test(libcpre_weights, libcpo_weights, alternative = "greater")
```

#### Result

Our p-value is 0.2203. We fail to reject $H_0$. We are unable to say if people became more liberal nor conservative as a result of this.

### Practical Significance

We are going to calculate Cohen's d to measure effect size. 

```{r q1_6}
calcPooledSd <- function(v1, v2) {
  ((length(v1)-1)*sd(v1)^2 + (length(v2)-1)*sd(v2)^2)/(length(v1) + length(v2) - 2)
}
calcCohensD <- function(v1, v2) {
  (mean(v1) - mean(v2))/calcPooledSd(v1, v2)
}

(d = calcCohensD(libcpre_weights, libcpo_weights))
```

The Cohen's d value for this case is `r d` which is $< 0.8$. There is no practical significance observed between $\mu_1$ and $\mu_2$.

## 2. Were Republican voters (examine variable pid\_x) older or younger (variable dem\_age\_r\_x), on the average, than Democratic voters in 2012?

```{r q2_1}
# Define bins for voter age histograms
bins_ages <- seq(from = 20, to = 100, by = 10)
bins_ages <- c(18, bins_ages)

# Extract Republican voters
rep_ages <- S$dem_age_r_x[grep("republican", S$pid_x, ignore.case = T)]
rep_ages <- rep_ages[is.numeric(rep_ages) & rep_ages >= 18]
rep_ages_n = length(rep_ages)
rep_age_mean = mean(rep_ages)
hist(rep_ages, main = "Republican Voter Age", xlab = "Age", breaks = bins_ages)

# Extract Democratic voters
dem_ages <- S$dem_age_r_x[grep("democrat", S$pid_x, ignore.case = T)]
dem_ages <- dem_ages[is.numeric(dem_ages) & dem_ages >= 18]
dem_ages_n = length(dem_ages)
dem_age_mean = mean(dem_ages)
hist(dem_ages, main = "Democratic Voter Age", xlab = "Age", breaks = bins_ages)
```

The average Republican voter age is `r rep_age_mean` which is in a little bit higher that that of the Democratic voeters' `r dem_age_mean`.

### Hypotheses

From this, we can formulate our hypotheses as below:

$\mu_1$ = Average Republican voter age\newline
$\mu_2$ = Average Democratic voter age

$$H_0: \mu_1 - \mu_2 = 0$$
$$H_1: \mu_1 - \mu_2 > 0$$

### Statistical Significance

#### Justificaition

Our n for Republicans is `r rep_ages_n` and for Democrats is `r dem_ages_n`. The distributions are fairly normal as seen above and our n's are sufficiently large. We can conduct a t-test in determining if $\mu_1$ is in fact greater than $\mu_2$.

```{r q2_2}
rs <- t.test(rep_ages, dem_ages, alternative = "greater")
p <- rs$p.value
```

#### Result

Our p-value is 0.0007. We can reject $H_0$ at a significance level of 0.05, 0.01 and 0.001.

### Practical Significance

We are going to calculate Cohen's D to measure effect size. 

```{r q2_3}
(d = calcCohensD(rep_ages, dem_ages))
```

The Cohen's d value for this case is `r d` which is $< 0.8$. There is no practical significance observed between $\mu_1$ and $\mu_2$.

## 3. Were Republican voters older than 51, on the average in 2012?

The average Republican voter is `r rep_age_mean`.

### Hypotheses

From this, we can formulate our hypotheses as below:

$$H_0: \mu = 51$$
$$H_1: \mu > 51$$

### Statistical Significance

#### Justificaition

Our n is `r rep_ages_n`. The distribution is fairly normal as seen in Q2 and our n is sufficiently large. We can conduct a t-test in determining if the average is in fact greater than 51 or not.

```{r q3_1}
rs <- t.test(rep_ages, alternative = "greater", mu = 51)
p <- rs$p.value
```

#### Result

Our p-value is `r p`. We fail to reject $H_0$ at a significance level of 0.05 or anything smaller.

### Practical Significance

We are going to calculate correlation r to measure effect size.

```{r q3_2}
computeEffectSizeCorrelationR <- function(t, df) {
  t/sqrt(t^2+df)
}
r = computeEffectSizeCorrelationR(rs$statistic, rep_ages_n-1)
```

The effect size correlation r for this case is `r r` which is $< 0.8$. There is no practical significance observed.

## 4. Were Republican voters more likely to shift their political preferences right or left (more conservative or more liberal), compared to Democratic voters during the 2012 election?

Let's compute how the average political standing changed for Republican voters.

```{r q4_1}
# Pre-2012 election Republican voters
libcpre_rep_weights <-compilePoliticalWeights(
  S$libcpre_self[grep("republican", S$pid_x, ignore.case = T)], 3, FALSE)
summary(libcpre_rep_weights)
mu_libcpre_rep = mean(libcpre_rep_weights)
hist(libcpre_rep_weights, main = "Pre-2012 Election Political Standing for Republican Voters", 
     xlab = "Political Standing", breaks = bins_standing)

# Post-2012 election Republican voters
libcpo_rep_weights <-compilePoliticalWeights(
  S$libcpo_self[grep("republican", S$pid_x, ignore.case = T)], 5, FALSE)
summary(libcpo_rep_weights)
mu_libcpo_rep = mean(libcpo_rep_weights)
hist(libcpo_rep_weights, main = "Post-2012 Election Political Standing for Republican Voters", 
     xlab = "Political Standing", breaks = bins_standing)

diff_lib_rep <- libcpo_rep_weights - libcpre_rep_weights
(diff_mu_lib_rep <- mean(diff_lib_rep))

(diff_mu_lib_rep = mu_libcpo_rep - mu_libcpre_rep)
```

The average political standing value post-2012 electioin is greater than pre-election by `r diff_mu_lib_rep`. This means that the Republican voters on average has become more conservative.

Now let's do the same for Democratic voters.

```{r q4_2}
# Pre-2012 election Democratic voters
libcpre_dem_weights <-compilePoliticalWeights(
  S$libcpre_self[grep("democrat", S$pid_x, ignore.case = T)], 3, FALSE)
summary(libcpre_dem_weights)
mu_libcpre_dem <- mean(libcpre_dem_weights)
hist(libcpre_dem_weights, main = "Pre-2012 Election Political Standing for Democratic Voters", 
     xlab = "Political Standing", breaks = bins_standing)

# Post-2012 election Democratic voters
libcpo_dem_weights <-compilePoliticalWeights(
  S$libcpo_self[grep("democrat", S$pid_x, ignore.case = T)], 5, FALSE)
summary(libcpo_dem_weights)
mu_libcpo_dem <- mean(libcpo_dem_weights)
hist(libcpo_dem_weights, main = "Post-2012 Election Political Standing for Democratic Voters", 
     xlab = "Political Standing", breaks = bins_standing)

diff_lib_dem <- libcpo_dem_weights - libcpre_dem_weights
(diff_mu_lib_dem <- mean(diff_lib_dem))
```

The average political standing value post-2012 electioin is smaller than pre-election by `r abs(diff_mu_lib_dem)`. This means that the Democratic voters on average has become more liberal.

### Hypotheses

From this, we can formulate our hypotheses as below:

$\mu_1$ = Average change in political standing for Republican voters\newline
$\mu_2$ = Average change in political standing for Democratic voters

$$H_0: \mu_1 - \mu_2 = 0$$
$$H_1: \mu_1 - \mu_2 > 0$$

### Statistical Significance

#### Justificaition

The Republican distribution is negatively skewed while the Democratic distribution is positively skewed. Therefore we should not use the t-test in determining statistical significance even though our n's are sufficiently large. Instead, we will use the nonpametric Wilcoxon rank-sum test.

```{r q4_3}
?wilcox.test
rs <- t.test(rep_ages, dem_ages, alternative = "greater")
p <- rs$p.value
```

#### Result

Our p-value is 0.0007. We can reject $H_0$ at a significance level of 0.05, 0.01 and 0.001.

### Practical Significance

We are going to calculate Cohen's D to measure effect size. 

```{r q2_3}
(d = calcCohensD(rep_ages, dem_ages))
```

## 5. Select a fifth question that you are interested in investigating.

Prepare a report addressing these questions.  A successful submission should include:

1. A brief introduction.

2. A suitable hypothesis test for each question above.

3. For each test, include:

    + A brief exploratory analysis targetted to check the assumptions needed for your test.

    + A justification for why the test is the most appropriate choice

    + An explanation of test results, including *BOTH* statistical significance and practical significance.

4. A brief conclusion with a few high-level takeaways.


Please limit your submission to 10 pages.  Be sure to submit both your pdf report as well as your source file.