---
title: 'Unit 5 Homework: Joint Distributions'
author: "Tako Hisada"
date: "10/7/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Unladen Swallows

```{r q1_setup}
library(MASS)
set.seed(898)
```
We know SD of Wingspan is 4. Since $V(X) = \sqrt(\sigma_X^2)$, $V(W) = 4^2 = 16$.

\[ V(V) = V[h(W)] = V[\frac{1}{2}w + U] = a^2*\sigma^2_W + V(U) = (\frac{1}{2})^2 * 4^2 + 1 = 5 \]

Now we want to compute $Cov(W, V)$.

\[ Cov(W, V) = Cov(W, \frac{1}{2}w + U) = Cov(W, \frac{1}{2}W) + Cov(W, U) = Cov(W, 1/2W) + 0 \]
\[ Cov(W, 1/2W) = E((W-E(W)(\frac{1}{2}W-E(\frac{1}{2}W)) = E((W-E(W)(\frac{1}{2}W-\frac{1}{2}E(W)) \]
\[ E((W-E(W)\frac{1}{2}(W-E(W)) = \frac{1}{2}E((W-E(W))^2) = \frac{1}{2}VAR(W) = \frac{1}{2}*16 = 8 \]

```{r q1}
# Wingspan
w = rnorm(100, mean = 10, sd = 4)
mu_w = 10
sd_w = 4
v_w = sd_w^2
a_w = 1/2

# Velocity
v = a_w * w + rnorm(100)
v_v = a_w^2 * sd_w^2 + 1

# Covariance
cov_w_v = 1/2*v_w

# Covariance-Variance Matrix
(Sigma = matrix(c(v_w, cov_w_v, cov_w_v, v_v), 2, 2))
```


# 2. Broken Rulers

## a. Conditional expectation of Y given X, E(Y|X)

X is uniformly distributed.
\[ f(Y|X) = \frac{1}{x} \]

\[ E(Y|X) = \int_{0}^X yf(Y|X)dy \]
\[ E(Y|X) = \int_{0}^X y*\frac{1}{x} = \frac{y^2}{x}|_{0}^{X} = \frac{x}{2} \]

## b. Unconditional expectation of Y

\[ E[Y] = E[E[Y|X]] \]

We know from a E[Y|X] = $\frac{x}{2}$

\[ E[Y] = E[ \frac{x}{2} ] = \frac{1}{2}E(X) \]

\[ E[X] = \int_0^1 xf(x)dx = \frac{x^2}{2}|_{0}^{1} = \frac{1}{2} \]

\[ E[Y] = \frac{1}{2}E[X] = \frac{1}{2}*\frac{1}{2} = \frac{1}{4} \]

## c. Compute E(XY)

\[ E[XY] = E[ E(XY|X) ] = E[ X*E(Y|X) ] = E[ X \frac{X}{2} ] = E[ \frac{X^2}{2} ] = \frac{1}{2}E[X^2] = \frac{1}{2}E[X^2] \]

\[ E[X^2] = E[h(X)] = \int_0^1 x^2f(x)dx = \int_0^1 x^2dx = \frac{x^3}{3}|_{0}^{1} = \frac{1}{3} \]

\[ E[XY] = \frac{1}{2}E[X^2] = \frac{1}{2}*\frac{1}{3} = \frac{1}{6} \]

## d. Compute Cov(X, Y)

\[ Cov(X, Y) = E(XY) - \mu_X\mu_Y = \frac{1}{6} - \frac{1}{2}*\frac{1}{4} = \frac{1}{24} \]

```{r q2d}
(q2d = 1/6-1/2*1/4)
```

# 3.  Great Time to Watch Async
## a. Total expected waiting time when you take the Caltrain each morning and each evening for 5 days in a row

```{r q3a-1}
i <- 0:5
ex = sum(1/6*i)
j <- 0:10
ey = sum(1/11*j)
exy = ex + ey
```
\[ E(X+Y) = E(X) + E(Y) \]
\[ E(X) = \sum_{x=0}^{5}x*p(x) = 0 * \frac{1}{6} + 1 * \frac{1}{6} + 2 * \frac{1}{6} + 3 * \frac{1}{6} + 4 * \frac{1}{6} + 5 * \frac{1}{6} = 2.5 \]
\[ E(Y) = \sum_{x=0}^{10}y*p(y) = 0 * \frac{1}{10} + 1 * \frac{1}{10} + 2 * \frac{1}{10} + 3 * \frac{1}{10} + 4 * \frac{1}{10} + 5 * \frac{1}{10} + 6 * \frac{1}{10} + 7 * \frac{1}{10} + 8 * \frac{1}{10} + 9 * \frac{1}{10} + 10 * \frac{1}{10} = 5 \]
\[ E(X+Y) = E(X) + E(Y) = 2.5 + 5 = 7.5 \]

Since we are being asked for the total expected waiting time when we took the Caltrain for 5 days in a row, we need to multiply this value by 5.

```{r q3a-2}
q3a = exy*5
```

\[ E(X+Y) = 7.5 * 5 = 37.5 \]

37.5 minutes

## b. Variance of the total waiting time

\[ V(X+Y) = V(X) + V(Y) + 2*COV(X+Y)\]
\[ V(X) = E(X^2) - E(X)^2 \]
\[ V(Y) = E(Y^2) - E(Y)^2 \]

```{r q3b_1}

compExUniform <- function(x) {
  sum = 0
  for(i in 0:x) {
    sum = sum + i * 1/(x+1)
  }
  sum
}

ex = compExUniform(5)
ey = compExUniform(10)

compExUniformSq <- function(x) {
  sum = 0
  for(i in 0:x) {
    sum = sum + i^2 * 1/(x+1)
  }
  sum
}
exx_sq <- compExUniformSq(5)
exy_sq <- compExUniformSq(10)

varx <- exx_sq - ex^2
vary <- exy_sq - ey^2
```

\[ Cov(X, Y) = E(XY) - \mu_{X}\mu_Y \]

```{r q3b_2}
compExXYUniform <- function(x, mux, y, muy) {
  sum = 0
  for(i in 0:x) {
    for(j in 0:y) {
      sum = sum + ((i-mux)*(j-muy))/((x+1)*(y+1))
    }
    #print(sprintf("END SUM: %f", sum))
  }
  sum
}
exxy <- compExXYUniform(5, 2.5, 10, 5)

cov_xy = exxy - ex*ey
```

Now we can compute $V(XY)$

```{r q2b_3}
varxy = varx + vary - 2*cov_xy
```

\[ 37\frac{11}{12} \]

However we need to multiple this by 5 as we are interested in finding out the variance of the total waiting time

```{r q2b_4}
q2b = 5^2 * varxy
```

\[ 947\frac{11}{12} \]

## c. Expected value of the difference between the total evening waiting time and the total morning waiting time over all 5 days

\[ E(X-Y) \]

```{r q3c}
i <- 0:5
ex = sum(1/6*i)
j <- 0:10
ey = sum(1/11*j)
exy_diff = abs(ex - ey)
```
\[ E(X-Y) = E(X) - E(Y) \]
\[ E(X) = \sum_{x=0}^{5}x*p(x) = 0 * \frac{1}{6} + 1 * \frac{1}{6} + 2 * \frac{1}{6} + 3 * \frac{1}{6} + 4 * \frac{1}{6} + 5 * \frac{1}{6} = 2.5 \]
\[ E(Y) = \sum_{x=0}^{10}y*p(y) = 0 * \frac{1}{10} + 1 * \frac{1}{10} + 2 * \frac{1}{10} + 3 * \frac{1}{10} + 4 * \frac{1}{10} + 5 * \frac{1}{10} + 6 * \frac{1}{10} + 7 * \frac{1}{10} + 8 * \frac{1}{10} + 9 * \frac{1}{10} + 10 * \frac{1}{10} = 5 \]
\[ E(X-Y) = |E(X) - E(Y)| = 2.5 \]

Since we are being asked for the total expected waiting time when we took the Caltrain for 5 days in a row, we need to multiply this value by 5.

```{r q3c_2}
q3c = exy_diff*5
```

\[ E(X+Y) = 2.5 * 5 = 12.5 \]

12.5 minutes

## d. Variance of the difference between the total evening time and the total morning waiting time over all 5 days

\[ V(X-Y) = V(X) + V(Y) \]
\[ V(X) = E(X^2) - E(X)^2 \]
\[ V(Y) = E(Y^2) - E(Y)^2 \]

```{r q3d_1}

compExUniform <- function(x) {
  sum = 0
  for(i in 0:x) {
    sum = sum + i * 1/(x+1)
  }
  sum
}

ex = compExUniform(5)
ey = compExUniform(10)

compExUniformSq <- function(x) {
  sum = 0
  for(i in 0:x) {
    sum = sum + i^2 * 1/(x+1)
  }
  sum
}
exx_sq <- compExUniformSq(5)
exy_sq <- compExUniformSq(10)

varx <- exx_sq - ex^2
vary <- exy_sq - ey^2
```

Now we can compute $V(XY)$

```{r q3d_2}
varxy = varx + vary
```

\[ 12\frac{11}{12} \]

However we need to multiple this by 5 as we are interested in finding out the variance of the total waiting time

```{r q3d_3}
q3d = 5^2 * varxy
```

\[ 322\frac{11}{12} \]

# 4. Maximizing Correlation

Show that if $Y = aX + b$ where X and Y are random variables and a != 0, corr(X, Y) = -1 or +1

\[ Corr(X, Y) = P_{X, Y} = \frac{Cov(X, Y)}{\sigma_{X}*\sigma_{Y}} \]
\[ Corr(X, aX+b) = P_{X, aX+b} = \frac{Cov(X, aX+b)}{\sigma_{X}*\sigma_{aX+b}} \]

Since adding a constant to random variables does not change their covariances or variances, we can say:

\[ \frac{Cov(X, aX)}{\sigma_{X}*\sigma_{aX}} \]

We can also move the constant a out of Cov and $\sigma$

\[ \frac{a*Cov(X, X)}{|a|*\sigma_{X}*\sigma_{X}} \]

We can cancel the constant a

\[ \frac{Cov(X, X)}{\sigma_{X}*\sigma_{X}} = \frac{Var(X)}{Var(X)} = 1 \]

In case a is a negative number, however, -1 remains in the numerator which makes Corr(X, Y) =  -1.
\[ \frac{-Cov(X, X)}{\sigma_{X}*\sigma_{X}} = \frac{Var(X)}{Var(X)} = -1 \]
```