---
title: 'Unit 3 Homework 2: Probability Theory'
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Gas Station Analytics

### (a) Probability that the next customer will request regular gas and fill the tank

Probability that the next customer use reulgar gas: 40%
Probability that will fill the tank: 30%

```{r 1-a}
(q1a <- 0.4 * 0.3)
```

12%

### (b) Probability that the next customer will fill the tank

#### Regular gas and fill the tank

Probability that the next customer use reulgar gas: 40%
Probability that will fill the tank: 30%

```{r 1-b-reg}
reg <- 0.4 * 0.3
```

#### Mid-grade gas and fill the tank

Probability that the next customer use mid-grade gas: 35%
Probability that will fill the tank: 60%

```{r 1-b-mid}
mid <- 0.35 * 0.6
```

#### Premium gas and fill the tank

Probability that the next customer use premium gas: 25%
Probability that will fill the tank: 50%

```{r 1-b-premium}
premium <- 0.25 * 0.5
```

```{r 1-b-sol}
(q1b <- reg + mid + premium)
```

45.5%

### (c) Conditional probability given the next customer fills the tank, they use regular gas

We know the probabilities of the next customer using regular gas and filling the tank and the total probability of the next customer filling the tank from (a) and (b).

```{r 1-c}
(q1c <- q1a/q1b)
```

26.4%

## 2. The Toy Bin

```{r q2-setup}
library(VennDiagram)
```

### (a) Area diagram representing events

```{r 2-a}
grid.newpage()
#draw.triple.venn(area1 = 1/2, area2 = 1/2, area3 = 1/3, n12 = 1/4, n13 = 1/6, n23 = 1/6, n123 = 1/12, category = c("Red", "Waterproof", "Cool"),  fill = c("red", "light blue", "green"), lty = ("blank"), alpha = rep(0.5, 1),  euler.d = TRUE, sep.dist = 0.03)
draw.quad.venn(area1 = 1/2, area2 = 1/2, area3 = 1/3, area4 = 1/6, n12 = 1/4, n13 = 1/6, n14 = 0, n23 = 1/6, n24 = 0, n34 = 0, n123 = 1/12, n1234 = 0, n124 = 0, n134 = 0, n234 = 0, category = c("Red", "Waterproof", "Cool", "Not Red, Waterproof nor Cool"),  fill = c("red", "light blue", "green", "yellow"), lty = ("blank"), alpha = rep(0.5, 1),  euler.d = TRUE, sep.dist = 0.03)
```

12%

\begin{figure}[h]
\includegraphics[width=0.5\textwidth, inner]
```{r test}
grid.newpage()
#draw.triple.venn(area1 = 1/2, area2 = 1/2, area3 = 1/3, n12 = 1/4, n13 = 1/6, n23 = 1/6, n123 = 1/12, category = c("Red", "Waterproof", "Cool"),  fill = c("red", "light blue", "green"), lty = ("blank"), alpha = rep(0.5, 1),  euler.d = TRUE, sep.dist = 0.03)
draw.quad.venn(area1 = 1/2, area2 = 1/2, area3 = 1/3, area4 = 1/6, n12 = 1/4, n13 = 1/6, n14 = 0, n23 = 1/6, n24 = 0, n34 = 0, n123 = 1/12, n1234 = 0, n124 = 0, n134 = 0, n234 = 0, category = c("Red", "Waterproof", "Cool", "Not Red, Waterproof nor Cool"),  fill = c("red", "light blue", "green", "yellow"), lty = ("blank"), alpha = rep(0.5, 1),  euler.d = TRUE, sep.dist = 0.03)
```
\caption{Caption}
\label{fig:figure2}
\end{figure}

### (b) 


### (c) 

