---
title: 'Unit 4 Homework: Random Variables'
author: "Tako Hisada"
date: "10/01/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Best Game in the Casino

### (a) How much do you get paid if the coin comes up heads 3 times?

\[
X =
\Bigg\{
  \begin{tabular}{cc}
    \$0 & 0 heads \\
    \$2 & 1 head \\
    \$4 & 2 heads \\
    \$x & 3 heads \\
  \end{tabular}
\]

\[
P(X) =
\Bigg\{
  \begin{tabular}{cc}
    1/8 & 0 heads \\
    1/4 & 1 head \\
    1/4 & 2 heads \\
    1/8 & 3 heads \\
  \end{tabular}
\]

```{r q1a,echo=FALSE}
ex = 6
payout = (ex - 0*1/8 - 2*1/4 - 4*1/4)*8
```

\[ E(X) = 0 * \frac{1}{8} + 2 * \frac{1}{4} + 4 * \frac{1}{4} + x * \frac{1}{8} = 6 \]
\[ \frac{x}{8} = 6 - \frac{2}{4} - \frac{4}{4} = \frac{18}{4} \]
\[ x = 36 \]

You would get paid $36 if the coin comes up heads 3 times.

### (b) Cumulative probability function for winning from the game

\[
  F(X) = P(X \leq x) =
  \Bigg\{
    \begin{tabular}{cc}
      1/8 & x = 0 \\
      3/8 & x = 1 \\
      5/8 & x = 2 \\
      1 & x = 3 \\
    \end{tabular} 
\]

## 2. Processing Pasta

### (a) Cumulative probability function of L

\[ F(L) = P(L \leq l) = \int_{-\infty}^{l} f(y)dy = \int_{0}^{l} \frac{y}{2}dy = \frac{y^2}{4} = \frac{l^2}{4} \]

### (b) Expected length of the pasta E(L)

\[
E(L) = \int_{-\infty}^{\infty} l*f(l)dl = \int_{0}^{2} l*\frac{l}{2} = \int_{0}^{2} \frac{l^2}{2} = \frac{l^3}{2*3} = \frac{l^3}{6} |_{0}^{2} = \frac{2^3}{6} - 0 = \frac{4}{3}
\]

## 3. The Warranty is Worth It

### (a) Expected payout from the contract, E(X) = E(g(T)) using the expression for the expectation of a function of a random variable

\[
f(x) = \Bigg\{
  \begin{tabular}{cc}
    1 0 $\leq$ x $\leq$ 1 \\
    0 Otherwise \\
  \end{tabular}
\]

\[
E(X) = E(g(T)) = \int_{-\infty}^{\infty} g(t)*f(x)dx = \int_{0}^{1} 100(1-t)^\frac{1}{2} * 1dx = 100 \int_{0}^{1} (1-t)^\frac{1}{2} * dx
\]

\[ = -100 * \frac{2}{3}(1-t)^\frac{3}{2} = -\frac{200}{3}(1-t)^\frac{3}{2}|_{0}^{1} = \frac{200}{3} \]

$66.67

### (b) E(X) another way, by first characterizing the random variable X.

#### i. Value for T that results in the payoff X = x 

\[ X = x \]

\[
F(X) = P(X \leq x) =
\Bigg\{
  \begin{tabular}{cc}
    0 & x<0 \\
    $\sqrt{100(1-t)}$ & 0 $\leq$ x $\leq$ 100 \\
  \end{tabular}
\]

\[ 100(1-t)^\frac{1}{2} = x \]
\[(1-t)^\frac{1}{2} = \frac{x}{100} \]
\[ \sqrt(1-t) = \frac{x}{100} \]
\[ 1-t = \frac{x^2}{10000}\]
\[ t = 1 - \frac{x^2}{10000} \]

#### ii. Condition for T that is equivalent to X $\leq$ x.

From i), we know
\[ t = 1 - \frac{x^2}{10000} \]

The T equivalent of X $\leq$ x would be
\[ T \geq 1 - \frac{x^2}{10000} \]

#### iii. Probability that X $\leq$ x using the condition above (cumulative probability function of X)

\[ F(Y) = P(X \leq x) = \int_0^{x} f(y)dy = \int_0^{x} 1-\frac{y^2}{10000}dy = x - \frac{x^3}{30000} \]

#### iv. Take a derivative to compute the probability density function for X

\[ f(x) = F'(X) = (x - \frac{x^3}{30000})dx = 1 - \frac{x^2}{10000} \]

#### v. Use the pdf of X to compute E(X)

\[ E(X) = \int_{-\infty}^{\infty} x*f(x)dx = \int_{0}^{100} x*(1-\frac{x^2}{10000}) dx \]
\[ = \frac{x^2}{2}-\frac{x^4}{40000} |_{0}^{100} = 10000 - \frac{1000000}{40000} = 2500 \]

This is clearly wrong..

## 4. The Baseline for Measuring Deviations

### (a) Expression for E(Y) and use properties of expectation 

\[ E(Y) = \int_{-\infty}^{\infty} (x-t)^2 f(x)dx = \frac{x^3*f(x)}{3} -x^2t * f(x) + t^2 * f(x) \]

### (b) Taking a partial derivative with respect to t, compute the value of t that minimizes E(Y)

\[ \frac{d}{dt}(\frac{x^3*f(x)}{3} -x^2t * f(x) + t^2 * f(x)) = t*f(x) - x^2*f(x) \]
\[ t*f(x) = x^2*f(x) \]
\[ t = x^2 \]

### (c) What is the value of E(Y) for this choice of t?

\[ E(Y) = \frac{x^3*f(x)}{3} -x^2t * f(x) + t^2 * f(x) = \frac{x^3*f(x)}{3} -x^4 * f(x) + x^4 * f(x) \]
